---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Reproducable Research Week 2 Assignment
### By Joe M

Note: Code chunks are set to show by default using: opts_chunk$set(echo=TRUE)
```{r set_Options,echo = FALSE}
  ## Set the behaviors of the R chucks to always show by default
  opts_chunk$set(echo=TRUE)
```

## To make sure the code is reproducable first we will verify the enviroment

Note: If this code is ran again, it is expected that the code is being ran from the working directory.
```{r set_Enviroment}
## Variables for this code - Since getWD is used here, it is expected that this code is being ran in the working directory
wd <- getwd()                                                                     ## Set Working location 
class_ZIP <- paste(wd, "activity.zip", sep = '/')                                 ## Create Zip File string
class_URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip" # Class zip location
class_CSV <- paste(wd, "activity.csv", sep = '/')                                 ## Create CSV File string

## Zip check - get if not there
if (!file.exists(class_ZIP)) {download.file(class_URL, class_ZIP)}                 ## If file does not exist, retrieve it

## CSV file check - Unzip crom class zip if not there
if (!file.exists(class_CSV)) {unzip(class_ZIP, overwrite = TRUE)}                 ## If file does not exist, retrieve it

'''

## Loading and preprocessing the data
### make sure the zip file exists -- If not retreive a new copy form the class page
wd <- getwd()                                                                     ## First get working Directory


if !file.exists


<a href="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip" rel="nofollow">Activity monitoring data</a>









## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
